Although the latest versions of both \easybuild{} and Lmod already provide
flexible support for hierarchical module naming schemes, further enhancements
to extend the capabilities and improve the user experience are planned.
Here we outline several promising possibilities.

The latest public version of \easybuild{} is still unaware of Lmod-specific
concepts that can be included in module files, like properties and families.
Additionally, other concepts including non-strict version specifications, e.g.,
\texttt{\small load(atleast("GCC","4.8"))}, in 
load statements can only be included in Lua module files.
Integrating these capabilities into \easybuild{} will
greatly enhance the value of both products, since it would allow HPC sites to
effortlessly maintain module families and properties.

Although the required support is available, a deeper module hierarchy that
better matches the toolchain concept in \easybuild{} is worth working out.
Usually only the \texttt{Compiler} and \texttt{MPI} hierarchy levels are used,
while \easybuild{} toolchains can also include math libraries. Extending
the two-level hierarchy to capture this level of flexibility is not straightforward
however; BLAS/LAPACK/FFT functionality can be provided by one single library (e.g.,
Intel MKL), or by multiple distinct libraries (e.g., OpenBLAS, LAPACK and FFTW).
This complicates the design of a deeper hierarchy that supports swapping one
math library for one or more alternatives.

Related to this is the more generic concept of a `multi-dimensional' module hierarchy,
also referred to as a module `matrix', which is currently an area of active research.
In such a multi-dimensional module hierarchy, the goal is that a module would
only become available when \emph{multiple} other modules are loaded together. This is
useful for software packages that provide support for multiple libraries that
incorporate some particular functionality, like the parallel
solver library PETSc which supports different LAPACK/FFT libraries. In such a module
hierarchy a module for PETSc, for example, would not be available until modules
for both e.g., OpenBLAS \emph{and} FFTW have been loaded. On top of this, the system
should also support swapping out the loaded modules for OpenBLAS and FFTW for one
single module providing equivalent functionality (e.g., an Intel MKL module), and
trigger a reload of the PETSc module. Supporting this would be particularly useful for
application developers and benchmarking studies.

Another improvement to \easybuild{} would be a dependency resolution
mechanism that is aware of subtoolchains. When the \easybuild{} framework checks the
availability of modules that match the required dependencies, it only uses the toolchain
selected for the software being installed (unless this is overridden specifically on a
per-dependency basis). This may require that modules which are, for example, only
dependent on a compiler are built and installed multiple times with different
toolchains, even though they are providing the same software builds. This yields
more modules than required and consumes additional disk space. Checking for
modules that were installed with a compatible subtoolchain (providing for example
only the compiler) would ensure that these tools only need to be installed once.

Mixing environment modules installed with different naming schemes leads to problems,
especially when both flat and hierarchical module naming schemes are in place.
As a result, migrating to a hierarchically organized set of modules can be painful.
To enable a smooth transition, a separate module tree with a different layout
would allow users to opt out of the old layout gradually.
This requires that \easybuild{} supports generating multiple module files for a
single software installation under different naming schemes. It should also be
possible to (re)generate only a module file for an existing software installation,
so that an alternative module stack can be populated without having to re-install the
software itself.

\ignore{
Furthermore, both \easybuild{} and Lmod need to be evaluated on, and possibly
ported to, special-purpose systems like Cray and IBM BlueGene. Preliminary
experiments are promising since they only reveal mostly minor issues on these
type of systems. On Cray systems for example, some care has to be taken in order
to be able to use the provided \texttt{ProgEnv} modules that wrap around the
different readily available compilers. One concept that is currently not well
supported yet by \easybuild{} is cross-compilation, in which software is built
for a particular type of system which is different than the one of which the build
is performed. This significantly complicates the running of tests, as is often
done by build configuration scripts.
\remark{mostly irrelevant to the paper topic}
}

%\markus{Future work: generate lua modulefiles from \easybuild{} including
%family specifications to prevent loading two compiler modules at the same
%time?}
